Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO

### 2016-02-29 Paper/paper.tex
stuff about multiple observations

\pagebreak
\newpage
\FloatBarrier
\section{Combining observations: looking forward with Advanced LIGO}\label{s1:multiple_observations}
% 
% 
\subsection{Multiple identical sources at low SNR}\label{s2:identical_multiple}
% 
% 
Here we show a crude calculation which tells us that approximately $8$ observations
of NSBH signals with $\rho\geq 10$ distributed uniformly in effective volume, would
give us similarly accurate a measurement of $\lambdans$ as would a single detection
with $\rho=50$. We follow the outline in Ref.~\cite{Markakis:2010mp}.

@@

There are two main caveats to this calculation, (i) it applies to identical sources, 
which is astrophysically next to impossible; (ii) the average measurement error 
$\sigma_{avg}$ has been defined as in the first equality in 
Eq.~\ref{eq:rmsSigmaIdenticalSources}.

% 
\subsection{Astrophysical source population}\label{s2:astro_multiple}
%

>>>>

\pagebreak
\newpage
\FloatBarrier
\section{Combining observations: looking forward with Advanced LIGO}\label{s1:multiple_observations}
% 
% 
\subsection{Multiple identical sources at low SNR}\label{s2:identical_multiple}
% 
% 
Here we show a crude calculation which tells us that approximately $8$ (or $16$) observations
of NSBH signals with $\rho\geq 10$ distributed uniformly in effective volume, would
give us similarly accurate a measurement of $\lambdans$ as would a single detection
with $\rho=50$ (or $\rho=70$).

@@

There are two main caveats to this rudimentary calculation, which was outlined in 
Ref.~\cite{Markakis:2010mp}, (i) it applies to identical sources,
which is astrophysically next to impossible to achieve exactly, but could perhaps happen 
approximately; (ii) the average measurement error $\sigma_{avg}$ has been defined as in 
the first equality in Eq.~\ref{eq:rmsSigmaIdenticalSources}.

% 
\subsection{Astrophysical source population}\label{s2:astro_multiple}
% 
Lets try a respectable Bayesian approach. Imagine we have $N$ stretches of data,
$d_1, d_2, \cdots, d_N$, each containing a single observation of an NSBH binary. Let $K$
denote our collective prior knowledge, except for expectations on binary parameters
themselves, which will enter the following calculation explicitly. Let the gravitational wave
signal from each binary be characterized by the parameters 
$\vec{\theta} = \{\mbh, \mns, \chibh, \chins, \rho\}$ and $\lambdans$. Note that we have folded
in the luminosity distance, inclination angle of the binary's orbital angular momentum with
respect to the plane of the detector, and its sky location angles into a single parameter $\rho$,
which is the SNR of the received signal. Note also that $\vec{\theta}$ does not include
$\lambdans$ as the latter would be treated independently. In addition, we assume
that all observed NS's have the same equation of state, that uniquely determines their
$\lambdans$ parameter. In other words, we assume that we know its mass perfectly.
% 
Using Bayes' theorem, the measured probability distribution function for the $\lambdans$
parameter is given as
\begin{equation}\label{eq:lambdaMultiple}
 p(\lambdans | d_1, d_2, \cdots, d_N; K) = p(\lambdans | K)^{\alpha}\prod_{i=1}^N p(\lambdans | d_i, K),
\end{equation}
where $p(\lambdans|K)$ is the prior expectation for $\lambdans$, and we have assumed
that all of the $N$ observations are independent. [Note: $\alpha$ will be calculated 
soon, it should be a function only of N.]. A priori, we will assume that no particular value
of $\lambdans$ is more preferred over another, and $\lambdans\leq 4000$, ie 
$$
p(\lambdans | K) = \mathrm{Rect}\left(\frac{\lambdans-2000}{4000}\right)
$$
% 
The second set of terms $p(\lambdans | d_i, K)$ are the inferred posterior distribution
functions for $\lambdans$ from \textit{each} observation independently, and are 
calculated by marginalizing over all other binary parameters, i.e.
\begin{equation}
 p(\lambdans | d_n, K) = \int \D\, \vec{\theta} p(\vec{\theta}, \lambdans | d_n, K).
\end{equation}
The joint inferred probability distribution of all parameters $\vec{\theta}\cup\{\lambdans\}$ 
is given by
\begin{equation}
 p(\vec{\theta}, \lambdans | d_n, K) = \dfrac{p(d_n|\vec{\theta}, \lambdans, K)\,p(\vec{\theta}, \lambdans | K)}{p(d_n|K)}.
\end{equation}
Here, $p(\vec{\theta}, \lambdans | K)$ is the prior probability for binary parameters of taking 
particular values. In this paper, we use a uniform prior on the individual masses within 
pre-chosen ranges, as well as a uniform prior on the spin of the black hole. We set neutron star's
spin identically $=0$, and its tidal deformability parameter $\lambdans$ to be equally likely
between $[0, 4000]$. The prior probability of obtaining a particular realization of data
$p(d_n|K)$ is absorbed into the overall normalization. Finally, the first term in the numerator
$p(d_n|\vec{\theta}, \lambdans, K)$ is the likelihood of obtaining the given stretch of 
data $d_n$ if we assume that a signal parameterized by $\vec{\theta}\cup\{\lambdans\}$ is buried
in it, i.e.
\begin{equation}\label{eq:likelihood}
 \LL(\vec{\theta}, \lambdans) \equiv p(\vec{\theta}, \lambdans | K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
\end{equation}
where $h\equiv h(\vec{\theta}, \lambdans)$ is our waveform model, and the detector noise
weighted inner-product $\langle a|b\rangle$ is defined as 
$$
\langle a|b\rangle \equiv 4\,\mathrm{Re}\left[\int_0^\infty \dfrac{\tilde{a}(f) \tilde(b)(f)^*}{S_n(|f|)}\,\D f\right],
$$
where $\tilde{a}$ means the Fourier transform of the finite time series $a(t)$, and 
$S_n(|f|)$ is the one-sided amplitude spectrum of detector noise. The definition of $\LL$ in 
Eq.~\ref{eq:likelihood} assumes that the instrument noise is colored Gaussian.

\prayush{%
As described in the previous sections, we have measured the $p(\lambdans | d_i, K)$ for a 
set of different binary parameters. Therefore, the problem of finding the cumulative measurement
constraint on $\lambdans$ using Eq.~\ref{eq:lambdaMultiple} reduces to being able to choose
an astrophysically motivated sample set of $N$ binaries over which the second term on the RHS 
can be computed.}

\prayush{%
Get posterior measurements for low SNRs, i.e. $\rho=\{10, 15, 20, 25, 30\}$. 
Fix black hole mass and spins. Compute the measurement certainty for $\lambdans$
as a function of the number of detections for a population astrophysically distributed 
in effective distance. Repeat for a few different combinations of black hole masses
and spins. This would be a sub-set of the overall detected population for NSBH binaries,
but what fraction of it would this comprise of is something that we can only probe with
LIGO itself. We do not know the real distribution of low-mass black holes in the field.
}
